---
title: "Practical Machine Learning Course Project"
author: "Mark Zollner"
date: "February 28, 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose 
This purpose of this project to create a machine learning model that reads in exercise data and predicts
the manner in which the participants completed the activity. There are six ways the exercise could have been done:
sitting-down, standing-up, standing, walking, or sitting. A random forest model, which can be used for classification, was chosen for this task.  

The data set is located here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv. The first step was 
to get rid of the columns of data that were empty, had "NA" entries, or "!DIV/0" errors. Then the data was divided into a
training (75%) and testing (25%) set.

```{r}
library(caret)
data <- read.csv("C:/pml-training.csv")
data <- subset(data, select = c(classe, gyros_forearm_x, gyros_forearm_y, gyros_forearm_z, accel_forearm_x, accel_forearm_y, accel_forearm_z, magnet_forearm_x, magnet_forearm_y, magnet_forearm_z, total_accel_forearm, gyros_dumbbell_x, gyros_dumbbell_y, gyros_dumbbell_z, accel_dumbbell_x, accel_dumbbell_y, accel_dumbbell_z, magnet_dumbbell_x, magnet_dumbbell_y,magnet_dumbbell_z, roll_forearm, pitch_forearm, yaw_forearm, total_accel_dumbbell, roll_dumbbell, pitch_dumbbell, yaw_dumbbell, gyros_arm_x, gyros_arm_y, gyros_arm_z, accel_arm_x, accel_arm_y, accel_arm_z, magnet_arm_x, magnet_arm_y, magnet_arm_z, gyros_belt_x, gyros_belt_y, gyros_belt_z, accel_belt_x, accel_belt_y, accel_belt_z, magnet_belt_x, magnet_belt_y, magnet_belt_z, roll_arm, pitch_arm, yaw_arm, total_accel_arm, roll_belt, pitch_belt, yaw_belt, total_accel_belt, num_window))
inTrain <- createDataPartition(y = data$classe, p = 0.75, list = FALSE)
training <- data[inTrain, ]
testing <- data[-inTrain, ]
```

Cross validation with ten folds was used for training the model. Ten was picked since this is a generally recommended value. Then the model was created with the "train" function. 
```{r}
train_control <- trainControl(method = "cv", number = 10)
rf_model <- train(classe ~ ., data = training, trControl = train_control, method = "rf")
```
The output from the model shows that it's very accurate, although this is just based on the training set. The classification errors for each of the exercises are very low, all below 1%.
```{r}
rf_model$results
```

## Determining the out of sample error
The confusion matrix shows that the out of sample error rate is also very low.
```{r}
pred <- predict(rf_model, testing)
confusionMatrix(pred, testing$classe)
```
## Conclusion
The random forest model created to predict the manner in which the exercise was completed is very accurate.


